---
title: "Effect Sizes"
author: "Chirag Patel"
date: "`r Sys.Date()`"
output: html_document
---


This new .Rmd incorporates the effect sizes comparing the top 10% vs. 1SD of exposure, and also has a new Manhattan Plot procedure.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(DBI)
library(ggsci)
library(DT)
library(ggrepel)
library(cowplot)
library(reactable)
library(gt)
library(broom)
library(ggridges)
```



```{r connect to db of summary stats}
#con <- DBI::dbConnect(RSQLite::SQLite(), dbname='../db/pe_summary_stats_08_2024.sqlite')
con <- DBI::dbConnect(RSQLite::SQLite(), dbname='../db/pe_summary_stats_01_2025.sqlite')
varnames <- tbl(con, "variable_names_epcf")
adjusted_meta <- tbl(con, "adjusted_meta_uwls") |> rename(term_name=expo_name)
mvr2 <- read_rds('../pe_summary_060623/mvr2.rds') |> mutate(mv = mve_rsq-base_rsq)
pe <- tbl(con, "pe") ## individual surveys
pe_r2 <- tbl(con, "rsq")
glanced <- tbl(con, "glanced")
glanced <- glanced |> left_join(pe_r2 |> select(exposure, phenotype, series, model_number,aggregate_base_model, rsq, adj.r2), by=c("exposure", "phenotype", "model_number", "aggregate_base_model", "series"))
variable_domain <- tbl(con, "variable_domain")
pe_overall <- tbl(con, "pe_overall")
r2_overall <- tbl(con, "rsq_overall")
glanced_overall <- tbl(con, "glanced_overall")
glanced_overall <- glanced_overall |> left_join(r2_overall |> select(exposure, phenotype, model_number,aggregate_base_model, rsq, adj.r2), by=c("exposure", "phenotype", "model_number", "aggregate_base_model"))

```

```{r}
expos <- pe |> filter(term %like% 'expo%') |> rename(evarname=exposure, pvarname=phenotype)
expos_wide <- expos |> pivot_wider(names_from = "model_number", values_from = c("estimate", "std.error", "statistic", "p.value")) 
glanced_wide <- glanced |> select(-c(adj.r2, df.residual, null.deviance, df.null, deviance)) |> pivot_wider(names_from=c("model_number", "aggregate_base_model"), values_from = c("rsq", "nobs", "AIC", "BIC")) |>  rename(evarname=exposure, pvarname=phenotype)
glanced_wide <- glanced_wide |> mutate(rsq_adjusted_base_diff=rsq_2_0-rsq_2_1, rsq_adjusted_diff = rsq_2_0-rsq_1_0) 
glanced_overall_wide <- glanced_overall |> select(-c(adj.r2, df.residual, null.deviance, df.null, deviance)) |> pivot_wider(names_from=c("model_number", "aggregate_base_model"), values_from = c("rsq", "nobs", "AIC", "BIC")) |>  rename(evarname=exposure, pvarname=phenotype)

expos_wide <- expos_wide |> left_join(glanced_wide |> select(-c(series, log_p, log_e, scaled_p, scaled_e)), by=c("evarname", "pvarname", "exposure_table_name", "phenotype_table_name"))
expos_wide <- expos_wide |> left_join(varnames, by=c("evarname"="Variable.Name", "exposure_table_name"="Data.File.Name"))
expos_wide <- expos_wide |> left_join(varnames |> select(Variable.Name, Data.File.Name, Variable.Description, Data.File.Description), 
                                      by=c("pvarname"="Variable.Name", "phenotype_table_name"="Data.File.Name"))

expos_wide <- expos_wide |> collect() |> select(-Use.Constraints) |> rename(e_data_file_desc=Data.File.Description.x, p_data_file_desc=Data.File.Description.y,e_variable_description=Variable.Description.x, p_variable_description=Variable.Description.y)

glanced_summary <- glanced_overall_wide |> mutate(total_n = nobs_2_0, rsq_adjusted_base_diff = rsq_2_0-rsq_2_1, rsq_adjusted_diff=rsq_2_0-rsq_1_0) |> 
  select(evarname, pvarname, rsq_adjusted_base_diff, rsq_adjusted_diff, total_n)

adjusted_meta_full <- adjusted_meta |> filter(model_number == 2)  |> left_join(glanced_summary, by=c("evarname", "pvarname")) |> collect() ## fully adjusted model

model_concordia <- adjusted_meta |> select(term_name, evarname, pvarname, estimate.uwls, model_number) |> group_by(term_name, evarname, pvarname) |> summarize(model_concordance = sum(sign(estimate.uwls))) |> collect()

model_concordia_12 <- adjusted_meta |> filter(model_number <= 2) |> select(term_name, evarname, pvarname, estimate.uwls) |> group_by(term_name, evarname, pvarname) |> summarize(model_concordance_1_2 = sum(sign(estimate.uwls))) |> collect()

adjusted_meta_full <- adjusted_meta_full |> left_join(model_concordia |> inner_join(model_concordia_12, by = c("term_name", "evarname", "pvarname")),  by = c("term_name", "evarname", "pvarname"))

p_variable_domain <- variable_domain |> filter(epcf == 'p') |> collect() |> group_by(Variable.Name) |> summarise(pvardesc=first(Variable.Description),pcategory=first(category),psubcategory=first(subcategory))
e_variable_domain <- variable_domain |> filter(epcf == 'e') |> collect() |> group_by(Variable.Name) |> summarise(evardesc=first(Variable.Description),ecategory=first(category),esubcategory=first(subcategory))

adjusted_meta_full <- adjusted_meta_full |> left_join(p_variable_domain, by=c("pvarname"="Variable.Name"))
adjusted_meta_full <- adjusted_meta_full |> left_join(e_variable_domain, by=c("evarname"="Variable.Name"))
expos_wide <- expos_wide |> left_join(p_variable_domain, by=c("pvarname"="Variable.Name"))
expos_wide <- expos_wide |> left_join(e_variable_domain, by=c("evarname"="Variable.Name"))
mvr2 <- mvr2 |> collect() |> left_join(p_variable_domain, by=c("pvarname"="Variable.Name"))

pe_overall_full <- pe_overall |> filter(model_number == 2, term %like% "expo%") |> select(term, estimate, std.error, statistic, p.value, exposure, phenotype) |> 
  rename(evarname = exposure, pvarname = phenotype, term_name=term) |> collect()

adjusted_meta_full <- adjusted_meta_full |> left_join(pe_overall_full, by=c("term_name", "evarname", "pvarname"), suffix=c("", "_overall"))

adjusted_meta_full <- adjusted_meta_full |> mutate(estimate_overall=ifelse(k.uwls == 1, estimate.uwls, estimate), 
                                                   p.value_overall=ifelse(k.uwls == 1, p.value.uwls, p.value),
                                                   std.error_overall=ifelse(k.uwls == 1, std.error.uwls, std.error), 
                                                   statistic_overall=ifelse(k.uwls == 1, statistic.uwls, statistic)
                                                   )
```

```{r filter exposure categories}
exposure_categories <- c("pollutant", "nutrients", "smoking", "infection")
adjusted_meta_full <- adjusted_meta_full |> filter(ecategory %in% exposure_categories)
expos_wide <- expos_wide |> filter(ecategory %in% exposure_categories)

```


```{r}
adjusted_meta_full <- adjusted_meta_full |> mutate(nobs = k.uwls)
n_obss <- sort(unique(adjusted_meta_full$nobs))

num_tests <- map_df(n_obss, function(n) {
  n_e <- adjusted_meta_full |> filter(nobs >= n) |> pull(evarname) |> unique() |> length()
  n_p <- adjusted_meta_full |> filter(nobs >= n) |> pull(pvarname) |> unique() |> length()
  nn <- nrow(adjusted_meta_full |> filter(nobs == n))
  tibble(n_expos=n_e, n_phenos=n_p, n_pxe=nn)
})

num_tests  |> mutate(n_surveys=n_obss) |> gt()



```

```{r}
# Keep number of surveys is greater than 2
adjusted_meta_2 <- adjusted_meta_full |> filter(nobs >= 2)
n_evars <- length(unique(adjusted_meta_2$evarname))
n_pvars <- length(unique(adjusted_meta_2$pvarname))
n_evars
n_pvars
```

Sample sizes within and across all surveys
```{r sample sizes}
sample_size_per_pair <- expos_wide |> filter(term == 'expo' | term== 'expo1') |> group_by(evarname, pvarname) |> summarize(total_n=sum(nobs_2_0), n_surveys=n(), median_n=median(nobs_2_0))

```

```{r}
adjusted_meta_2 <- adjusted_meta_2 |> filter(!is.na(estimate_overall), !is.na(p.value_overall))
adjusted_meta_2 <- adjusted_meta_2 |> ungroup() |>  mutate(pval_BY=p.adjust(p.value_overall, method="BY"), pvalue_bonferroni=p.adjust(p.value_overall, method="bonferroni"))
adjusted_meta_2 <- adjusted_meta_2 |> mutate(sig_levels = case_when(
  pvalue_bonferroni < 0.05 ~ 'Bonf.<0.05',
  pval_BY < 0.05 ~ 'BY<0.05',
  TRUE ~ '> BY & Bonf.'
))

```

## Sample for LLM 
```{r}
#set.seed(123)
#adjusted_meta_2 |> select(sig_levels, estimate, evarname, pvarname, evardesc, pvardesc) |> group_by(sig_levels) |> slice_sample(n=500) |> write_csv("../exwas_sample.csv")


```


```{r rename subcategories}

adjusted_meta_2 <- adjusted_meta_2 |> mutate(
  pnewsubcategory =  case_when(
    psubcategory == 'injury' ~ "inflammation",
    psubcategory == 'liver/kidney' ~ "kidney",
    psubcategory == 'immunity' ~ "liver",
    psubcategory == 'exhaled NO' ~ "inflammation",
    psubcategory == 'lung function' ~ "lung",
    psubcategory == 'nutritional status' ~ "nutrition",
    TRUE ~ psubcategory
  )
)

adjusted_meta_2 <- adjusted_meta_2 |> mutate(pnewsubcategory = ifelse(is.na(pnewsubcategory), pcategory, pnewsubcategory))
adjusted_meta_2 <- adjusted_meta_2 |> mutate(enewsubcategory = ifelse(ecategory == 'infection', 'infection', esubcategory))
adjusted_meta_2 <- adjusted_meta_2 |> mutate(enewsubcategory = ifelse(enewsubcategory == 'pyrethoid', 'pyrethroid', enewsubcategory)) 
adjusted_meta_2 <- adjusted_meta_2 |> mutate(enewsubcategory = ifelse(enewsubcategory == 'amine' | enewsubcategory == 'amide', 'amine/amide', enewsubcategory))
adjusted_meta_2 <- adjusted_meta_2 |> mutate(enewsubcategory = ifelse(enewsubcategory == 'deet', 'amine/amide', enewsubcategory))

```

```{r}
## phenotype-specific correction
adjusted_meta_2 <- adjusted_meta_2 |> group_by(pvarname) |> mutate(pval_BY_pheno=p.adjust(p.value_overall, method="BY")) |> ungroup()
```

```{r estimate check}
adjusted_meta_2 |> filter(evarname == 'LBXBPB', pvarname == 'LBXRDW') |> select(p.value_overall, p.value.uwls, estimate.uwls, estimate_overall)

p <- ggplot(adjusted_meta_2, aes(estimate_overall, estimate.uwls))
p <- p + geom_point() + facet_grid(~I(p.value_overall < 1e-5))
adjusted_meta_2 |> mutate(estimate_match = sign(estimate_overall) == sign(estimate.uwls)) |>  group_by(sig_levels) |> count(estimate_match)
adjusted_meta_2 |> mutate(estimate_match = sign(estimate_overall) == sign(estimate.uwls)) |> filter(!estimate_match) |> filter(sig_levels == 'Bonf.<0.05') |> arrange(p.value_overall)

adjusted_meta_2 |> mutate(estimate_match = sign(estimate_overall) == sign(estimate.uwls)) |> filter(!estimate_match) |> filter(sig_levels == 'Bonf.<0.05') |> group_by(pvarname) |> count() |> arrange(-n)

```




```{r}
con_q  <- DBI::dbConnect(RSQLite::SQLite(), dbname='../db/pe_summary_stats_percentile_scale_p_01_2025_v2.sqlite')
pe_overall_q <- tbl(con_q, "pe_overall")
pe_quantile <- pe_overall_q |> filter( model_number == 2) |> filter(term == 'expoq2' | term =='expoq3' | term == 'expoq4' | term=='expoq5') |> collect()
dbDisconnect(con_q)


pe_quantile <- pe_quantile |> rename(evarname=exposure, pvarname=phenotype)
pe_quantile <- pe_quantile |> mutate(exposure_level = case_when(
  term == 'expoq2' ~ 'P[25-50)',
  term == 'expoq3' ~ 'P[50-75)',
  term == 'expoq4' ~ 'P[75-90)',
  term == 'expoq5' ~ 'P[90-100)',
  TRUE ~ NA_character_
))


linear_res <- adjusted_meta_2 |> filter(term_name == 'expo') |> select(evarname, pvarname, estimate_overall, sig_levels, ecategory, esubcategory, pcategory, psubcategory, enewsubcategory, pnewsubcategory)

pe_quantile_all <- pe_quantile |> left_join(linear_res, by=c("evarname", "pvarname"))
pe_quantile_top <- pe_quantile_all |> filter(sig_levels == 'Bonf.<0.05')

p <- ggplot(pe_quantile_all |> filter(!is.na(sig_levels)) |> mutate(sig_levels = factor(sig_levels, levels=c("> BY & Bonf.", "BY<0.05", "Bonf.<0.05"))) ,  aes(y=exposure_level, x=I(abs(estimate)), fill=sig_levels))
p <- p + geom_boxplot()+ scale_x_continuous(limits=c(0, .7))
p <- p + xlab("Absolute Value of Effect Size [1SD of Phenotype]") + ylab("Quantile of Exposure") + scale_fill_nejm()
p1 <- p + theme_bw() + theme(legend.position = "bottom")

#ggsave("./paper figs/EffectSize.pdf", p1, width=5, height=5)

keep_p <- c("anthropometric", "blood", "blood pressure", "bone", "electrolyte", "inflammation", "kidney", "liver", "hormone", "metabolic", "nutrition", "dexa")

p <- ggplot(pe_quantile_top |> filter(pnewsubcategory %in% keep_p) , aes(y=exposure_level, x=I(abs(estimate)), fill=pnewsubcategory))
p <- p + geom_boxplot()+ scale_x_continuous(limits=c(0, .7)) 
p <- p + xlab("Absolute value of Effect Size [1SD of Phenotype]") + ylab("Quantile of Exposure")
p2 <- p + theme_bw() + theme(legend.position = "none")

enewsubcategory_labels <- c("DI","OR","DB","VO","PP","HM","HY","DA","PH","AA","SB","PT","PO","IN", "SU", "SM")
names(enewsubcategory_labels) <- c("dietary interview","organochlorine","dietary biomarker","VOC","priority pesticide","heavy metals","hydrocarbon","diakyl","phenols","amine/amide","smoking biomarker","phthalates","polyfluoro","infection","supplements","smoking behavior" )

p <- ggplot(pe_quantile_top |> filter(pnewsubcategory %in% keep_p) , aes(x=exposure_level, y=I(abs(estimate)), fill=enewsubcategory))
p <- p + geom_boxplot() + scale_y_continuous(limits=c(0, .7)) 
p <- p + ylab("Effect Size [1SD of Phenotype]") + xlab("Quantile of Exposure")
p <- p + theme_bw() + 
  facet_grid(enewsubcategory~pnewsubcategory, labeller = labeller(enewsubcategory = enewsubcategory_labels)) + 
  coord_flip() + theme(legend.position = "bottom",axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=1))


ggsave("./paper figs/EffectSize_matrix.pdf", p, width=8.5, height=10)
```




# Examples on the original units

- HABC% (LBXGH)
- Fasting glucose (LBXGLU)
- LDL (LBDLDL)
- Triglycerides (LBXSTR)
- Serum Creatinine (LBXSCR)
- CRP (LBXCRP)
- Thyroid (LBXTSH1)
- AST/ALT (LBXSATSI/LBXSASSI)
- Blood Pressure (MSYSTOLIC)


```{r}
con_q  <- DBI::dbConnect(RSQLite::SQLite(), dbname='../db/pe_summary_stats_percentile_noscale_p_01_2025_v2.sqlite')
pe_overall_q <- tbl(con_q, "pe_overall")
pe_quantile_ns <- pe_overall_q |> filter( model_number == 2) |> filter(term == 'expoq2' | term =='expoq3' | term == 'expoq4' | term=='expoq5') |> collect()
dbDisconnect(con_q)


pe_quantile_ns <- pe_quantile_ns |> rename(evarname=exposure, pvarname=phenotype)
pe_quantile_ns <- pe_quantile_ns |> mutate(exposure_level = case_when(
  term == 'expoq2' ~ 'P[25-50)',
  term == 'expoq3' ~ 'P[50-75)',
  term == 'expoq4' ~ 'P[75-90)',
  term == 'expoq5' ~ 'P[90-100)',
  TRUE ~ NA_character_
))


linear_res <- adjusted_meta_2 |> filter(term_name == 'expo') |> select(evarname, pvarname, estimate_overall, estimate.uwls, sig_levels, ecategory, esubcategory, pcategory, psubcategory, enewsubcategory, pnewsubcategory, evardesc, pvardesc, model_concordance, model_concordance_1_2) 

pe_quantile_ns <- pe_quantile_ns |> left_join(linear_res, by=c("evarname", "pvarname"))
pe_quantile_ns_top <- pe_quantile_ns |> filter(sig_levels == 'Bonf.<0.05', abs(model_concordance_1_2) == 2, evarname != "LBXHCT", sign(estimate_overall) == sign(estimate.uwls))


pe_quantile_ns |> group_by(sig_levels, abs(model_concordance_1_2==2), sign(estimate_overall) == sign(estimate.uwls)) |> count()
```

```{r}
## top ones
plot_effect_size_top_n <- function(quantile_assoc_data, pname, top_number = 7, show_legend="bottom", ylabel="Effect Size vs. 0-25% Exposure", xlabel="") {
  top_evars <- quantile_assoc_data |> filter(pvarname == pname, term=='expoq5') |> slice_min(n=top_number, order_by = p.value) |> select (evarname)
  top_evars <-  top_evars |> left_join(quantile_assoc_data |> filter(pvarname==pname))
  p <- ggplot(top_evars, aes(exposure_level, estimate, group=evardesc, color=evardesc))
  p <- p + geom_point() + geom_line() 
  p <- p + geom_errorbar(aes(ymin=estimate-std.error*1.96, ymax=estimate+std.error*1.96), width=.1)
  p <- p + geom_line()  + ylab(ylabel) + xlab(xlabel) + scale_color_tron()
  p <- p + theme_bw() + theme(legend.position=show_legend)
  p
}

plot_effect_sizes <- function(pname) {
  p <- ggplot(pe_quantile_ns_top, aes(exposure_level, estimate, group = evardesc))
  p <- p + geom_point() + geom_line()  + ylab("Effect Size [XXX vs. lowest group]")
  p <- p + geom_line()  + ylab("Effect Size [ lowest group]") + xlab("Exposure Group")
  p <- p + theme_bw()
  p
}

plot_effect_size_top_n(pe_quantile_ns_top, "LBXGLU", top_number = 6) + ggtitle("Glucose") # show
plot_effect_size_top_n(pe_quantile_ns_top, "LBXGH", top_number = 6) + ggtitle("Hemoglobin A1C (%)") # show
plot_effect_size_top_n(pe_quantile_ns_top, "LBXCRP", top_number = 6) + ggtitle("Serum C-Reactive Protein (mg/dL)") # show
plot_effect_size_top_n(pe_quantile_ns_top, "BMXHT", top_number = 8)  + ggtitle("Height (cm)")
plot_effect_size_top_n(pe_quantile_ns_top, "BMXBMI", top_number = 6) + ggtitle("Body Mass Index (kg/m2)") # show
plot_effect_size_top_n(pe_quantile_ns_top, "LBDLDL", top_number = 6) + ggtitle("LDL Cholesterol (mg/dL)")
plot_effect_size_top_n(pe_quantile_ns_top, "LBXSCR", top_number = 6) + ggtitle("Serum Creatinine (mg/dL)") # show 
plot_effect_size_top_n(pe_quantile_ns_top, "LBXTSH1", top_number = 6) + ggtitle("Thyroid Stimulating Hormone (mcIU/mL)")
plot_effect_size_top_n(pe_quantile_ns_top, "LBXSATSI", top_number = 6) + ggtitle("Aspartate Aminotransferase (U/L)")
plot_effect_size_top_n(pe_quantile_ns_top, "LBXSASSI", top_number = 6) + ggtitle(" Alanine Aminotransferase (U/L)")
q # show 
plot_effect_size_top_n(pe_quantile_ns_top, "LBXWBCSI", top_number = 6) + ggtitle("White Blood Cells (1000 cells/uL)") # show
plot_effect_size_top_n(pe_quantile_ns_top, "SPXNFEV1", top_number = 8) + ggtitle("Forced Expiratory Volume in 1 Second (mL)") # show
plot_effect_size_top_n(pe_quantile_ns_top, "LBXEOPCT", top_number = 6)
plot_effect_size_top_n(pe_quantile_ns_top, "LBXHCT", top_number = 6)
plot_effect_size_top_n(pe_quantile_ns_top, "LBXRDW", top_number = 6) + ggtitle("Red cell Distribution Width (%)") # 
```




# cdf of the R2

```{r cdfs}

adjusted_meta_2 <- adjusted_meta_2 |> mutate(vartype = ifelse(term_name == 'expo', 'continuous', 'categorical'))
p <- ggplot(adjusted_meta_2 |> filter(vartype =='categorical'), aes(abs(estimate), color=sig_levels))
p <- p + stat_ecdf() + scale_x_continuous(limits=c(0, .25))
p <- p + xlab("abs(estimate)") + ylab("percentile")  + theme(legend.position="bottom") + scale_color_npg()
p

p <- ggplot(adjusted_meta_2 |> filter(vartype =='continuous'), aes(abs(estimate), color=sig_levels))
p <- p + stat_ecdf() + scale_x_continuous(limits=c(0, .25))
p <- p + xlab("abs(estimate)") + ylab("percentile") + theme(legend.position="bottom") + scale_color_npg()
p


p <- ggplot(adjusted_meta_2, aes(abs(estimate), color=sig_levels))
p <- p + stat_ecdf() + scale_x_continuous(limits=c(0, .25))
p <- p + xlab("abs(estimate)") + ylab("percentile") + theme(legend.position="bottom") + scale_color_npg()
p

ecdf_for_sig <- adjusted_meta_2 |> filter(sig_levels == 'Bonf.<0.05') |> pull(rsq_adjusted_base_diff) |> ecdf()
ecdf_for_non_sig <- adjusted_meta_2 |> filter(sig_levels == '> BY & Bonf.') |> pull(rsq_adjusted_base_diff) |> ecdf()

p_percent_sig <- adjusted_meta_2 |> group_by(pnewsubcategory, sig_levels) |> count() |> pivot_wider(names_from = sig_levels, values_from = n) |> mutate(total_tests = rowSums(across(where(is.numeric)), na.rm = TRUE)) |> 
  mutate(percent_non_sig = `> BY & Bonf.`/total_tests, percent_bonf = `Bonf.<0.05`/total_tests, percent_BY=`BY<0.05`/total_tests) |> 
  select(pnewsubcategory, total_tests, percent_non_sig, percent_bonf, percent_BY) |> ungroup()

p_effect_sizes_per <- adjusted_meta_2 |> filter(sig_levels == 'Bonf.<0.05') |> group_by(pnewsubcategory) |> summarize(median_pvalue=median(p.value), median_r2=median((rsq_adjusted_base_diff))) |> left_join(p_percent_sig)

p_effect_sizes_per <- p_effect_sizes_per |> mutate(q = ecdf_for_sig(median_r2), sig_levels ='Bonf.<0.05')
p_effect_sizes_per <- p_effect_sizes_per |> mutate(p_cat = pnewsubcategory)



p <- ggplot(adjusted_meta_2 |> mutate(sig_levels = factor(sig_levels, levels=c("> BY & Bonf.", "BY<0.05", "Bonf.<0.05"))), aes(rsq_adjusted_base_diff, color=sig_levels))
p <- p + stat_ecdf() + scale_x_continuous(limits=c(0, .05)) +scale_color_nejm()
p <- p + geom_point(data=p_effect_sizes_per, aes(x=median_r2, y = q, color=sig_levels)) 
p <- p + geom_text_repel(data=p_effect_sizes_per, aes(x=median_r2, y = q, color=sig_levels, label=p_cat)) 
p <- p + xlab("R^2 (adjusted-base model)") + ylab("Percentile") 
p <- p + theme_bw() + theme(legend.position="bottom") 
p
ggsave("./paper figs/r2_CDF.pdf", p, width=5, height=5)



```



Compare the pvalues from the two different approaches

```{r}

pval_compare <- pe_quantile |> filter(term == 'expoq5') |> select(evarname, pvarname, estimate, p.value) |> right_join(adjusted_meta_2, by=c("evarname", "pvarname"))

ggplot(pval_compare, aes(-log10(p.value.x), -log10(p.value_overall))) + geom_point()
```




