---
title: "Attaining Residuals from the Regression"
output: html_document
---


```{r setup, include=FALSE}
# Set global chunk options: show code by default
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This vignette shows how to extract residuals from the fitted models and plot the fitted vs. residual values to visualize presence of heteroskedascity. 

# 1.  Load Libraries


```{r load}
library(tidyverse) # for data‐manipulation (dplyr, %\>%, etc.)
library(DBI) # for connecting to an SQLite (or other) database
library(survey) # loaded by nhanespewas for complex‐survey regressions
library(nhanespewas) # our package for PE associations

```


# 2.  Connect to the NHANES PE Database

We assume that you have already downloaded the SQLite database
containing all variables (1999–2017) formatted for nhanespewas. Adjust
the path_to_db to match your local file.

```{r connect}
# Replace './db/nhanes_031725.sqlite' with your actual file path
path_to_db <- "../db/nhanes_031725.sqlite"
# Establish a connection to the NHANES PE database
con <- connect_pewas_data(path_to_db)

```

```{r summary stats loading}
load('adjusted_meta_2.Rdata')

```




# Example 1: Hemoglobin (P) vs. Cadmium (E)



```{r}
varnames |> filter(Variable.Name == "LBXBCD") 
varnames |> filter(Variable.Name == "LBXHGB")
```

## 2. Run PE association: Hemoglobin (outcome) \~ Cadmium (exposure) and plot the fitted vs. the residuals

 - adjustment_models: built‐in set of covariates (age, sex, race_ethnicity, etc.)

 - scale_p = TRUE: z‐score the phenotype (LBXHGB) before regression

 - logxform_p = FALSE: do NOT log‐transform the phenotype

 - scale_type = 1:  scaled phenotype by SD
 
 - save_svymodel = T: this will save the model to the workspace to extract the residuals.

This association had an very small pvalue and measured in a large sample.


Centered around a mean of zero with little evidence of non-constant variance.

```{r}
hgb_bcd <- pe_flex_adjust("LBXHGB","LBXBCD", adjustment_variables   = adjustment_models, con = con, scale_p = TRUE, logxform_p = FALSE, scale_type = 1, save_svymodel=T ) # saving the model
fitted_residual <- tibble(res=resid(hgb_bcd$models[[2]]$model), fitted=fitted(hgb_bcd$models[[2]]$model))
p <- ggplot(fitted_residual, aes(fitted,res)) # unweighted residuals
p <- p + geom_point(alpha=.05)  + theme_bw()
p
```
# Example 2: Glucose and beta-carotene

```{r}
glu_bec <- pe_flex_adjust("LBXGLU","LBXBEC", adjustment_variables   = adjustment_models, con = con, scale_p = TRUE, logxform_p = FALSE, scale_type = 1, save_svymodel=T ) # saving the model
fitted_residual <- tibble(res=resid(glu_bec$models[[2]]$model), fitted=fitted(glu_bec$models[[2]]$model))
p <- ggplot(fitted_residual, aes(fitted,res))
p <- p + geom_point(alpha=0.2) + theme_bw()
p
```


```{r}
dbDisconnect(con)
```

