---
title: "PxE"
author: "Chirag Patel"
date: "`r Sys.Date()`"
output: html_document
---

# Systematic Associations Between Phenotypes and Environmental Exposures: Benchmarking Exposomic Research
## NHANES

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r start anew}
if(exists("con")) {
  dbDisconnect(con)
  remove(list=ls())
}

```

```{r}
library(tidyverse)
library(DBI)
library(ggsci)
library(DT)
library(ggrepel)
library(cowplot)
library(reactable)
library(gt)
library(broom)


```

```{r combine within survey and glanced stats}
ep_assoc_summary_across_models <- function(summary_stats, glanced_stats) {
  summary_stats_wide <- summary_stats |> pivot_wider(names_from = "model_type", values_from = c("estimate", "std.error", "statistic", "p.value")) 
  summary_stats_wide <- summary_stats_wide |> mutate(estimate_diff = estimate_adjusted-estimate_unadjusted)
  adj_vs_base <- glanced |> select(-c(adj.r2, df.residual, null.deviance, df.null, deviance)) |> pivot_wider(names_from=model_type, values_from = c("rsq", "nobs", "AIC", "BIC"))
  adj_vs_base <- adj_vs_base |> mutate(rsq_adjusted_base_diff=rsq_adjusted-rsq_base, rsq_adjusted_diff = rsq_adjusted-rsq_unadjusted)
  summary_stats_wide |> left_join(adj_vs_base, by=c("evarname", "pvarname", "exposure_table_name", "phenotype_table_name"))
}

remove_units_from_string <- function(vardesc) {
  gsub("\\(.*\\)$","", vardesc)
}

```

```{r connect to db}
con <- DBI::dbConnect(RSQLite::SQLite(), dbname='../data/pe_summary_stats.sqlite')
varnames <- tbl(con, "variable_names_epcf")
adjusted_meta <- tbl(con, "adjusted_meta")
unadjusted_meta <- tbl(con, "unadjusted_meta")
adjusted_meta <- adjusted_meta |> left_join(unadjusted_meta |> select(evarname, pvarname, expo_name, vartype, estimate, p.value) |> rename(estimate_unadjusted=estimate, p.value_unadjusted=p.value), by=c("evarname", "pvarname", "expo_name", "vartype"))
mvr2 <- tbl(con, 'mvr2') |> mutate(mv = mve_rsq-base_rsq)
pe <- tbl(con, "pe")
glanced <- tbl(con, "glanced")
variable_domain <- tbl(con, "variable_domain")
```

```{r merge with group annotations}
expos <- pe |> filter(term %like% 'expo%')
expos_wide <- ep_assoc_summary_across_models(expos, glanced)
expos_wide <- expos_wide |> left_join(varnames, by=c("evarname"="Variable.Name", "exposure_table_name"="Data.File.Name"))
expos_wide <- expos_wide |> left_join(varnames |> select(Variable.Name, Data.File.Name, Variable.Description, Data.File.Description), 
                                      by=c("pvarname"="Variable.Name", "phenotype_table_name"="Data.File.Name"))

expos_wide <- expos_wide |> select(-Use.Constraints) |> rename(e_data_file_desc=Data.File.Description.x, p_data_file_desc=Data.File.Description.y,
                                                               e_variable_description=Variable.Description.x, 
                                                               p_variable_description=Variable.Description.y
                                                               )
expos_wide <- expos_wide |> collect()
#expos_wide_summary <- expos_wide |> filter(term == 'expo' | term == 'expo1') |> group_by(evarname, pvarname) |> summarize(mean_adjusted_base_r2_diff = mean(rsq_adjusted_base_diff), mean_unadjusted_r2_diff=mean(rsq_adjusted_diff), total_n = sum(nobs_adjusted)) |> ungroup()

expos_wide_summary <- expos_wide |> filter(term == 'expo' | term == 'expo1' | term == 'expo2') |> group_by(evarname, pvarname) |> summarize(mean_adjusted_base_r2_diff = mean(rsq_adjusted_base_diff), mean_unadjusted_r2_diff=mean(rsq_adjusted_diff), total_n = sum(nobs_adjusted)) |> ungroup()

adjusted_meta <- adjusted_meta |> collect() |> left_join(expos_wide_summary, by=c("evarname", "pvarname"))


p_variable_domain <- variable_domain |> filter(epcf == 'p') |> collect() |> group_by(Variable.Name) |> summarise(pvardesc=first(Variable.Description),pcategory=first(category),psubcategory=first(subcategory))
e_variable_domain <- variable_domain |> filter(epcf == 'e') |> collect() |> group_by(Variable.Name) |> summarise(evardesc=first(Variable.Description),ecategory=first(category),esubcategory=first(subcategory))

adjusted_meta <- adjusted_meta |> left_join(p_variable_domain, by=c("pvarname"="Variable.Name"))
adjusted_meta <- adjusted_meta |> left_join(e_variable_domain, by=c("evarname"="Variable.Name"))

expos_wide <- expos_wide |> left_join(p_variable_domain, by=c("pvarname"="Variable.Name"))
expos_wide <- expos_wide |> left_join(e_variable_domain, by=c("evarname"="Variable.Name"))

```

# Number of unique exposures and phenotypes

```{r}
num_e <- length(unique(adjusted_meta$evarname))
num_p <- length(unique(adjusted_meta$pvarname))
num_e
num_p

num_e * num_p

```

# Number exposures and phenotypes and associations in X number of surveys

```{r}
#adjusted_meta <- adjusted_meta |> unnest(glanced) |> unnest(tidied)
n_obss <- sort(unique(adjusted_meta$nobs))

num_tests <- map_df(n_obss, function(n) {
  n_e <- adjusted_meta |> filter(nobs == n) |> pull(evarname) |> unique() |> length()
  n_p <- adjusted_meta |> filter(nobs == n) |> pull(pvarname) |> unique() |> length()
  nn <- nrow(adjusted_meta |> filter(nobs == n))
  tibble(n_expos=n_e, n_phenos=n_p, n_pxe=nn)
})

num_tests  |> mutate(n_surveys=n_obss) |> gt()

```

# Keep number of surveys is greater than 2

```{r}
adjusted_meta_2 <- adjusted_meta |> filter(nobs >= 2)
n_evars <- length(unique(adjusted_meta_2$evarname))
n_pvars <- length(unique(adjusted_meta_2$pvarname))
```

# Sample sizes within and across all surveys

```{r sample sizes}
sample_size_per_pair <- expos_wide |> filter(term == 'expo' | term== 'expo1') |> group_by(evarname, pvarname) |> summarize(total_n=sum(nobs_adjusted), n_surveys=n(), median_n=median(nobs_adjusted))
```

# Summary of the summary stats

```{r fig.height=5, fig.width=10}
adjusted_meta_2 <- adjusted_meta_2 |> ungroup() |>  mutate(pval_BY=p.adjust(p.value, method="BY"), pvalue_bonferroni=p.adjust(p.value, method="bonferroni"))
adjusted_meta_2 <- adjusted_meta_2 |> mutate(sig_levels = case_when(
  pvalue_bonferroni < 0.05 ~ 'Bonf.<0.05',
  pval_BY < 0.05 ~ 'BY<0.05',
  TRUE ~ '> BY & Bonf.'
))

bonf_thresh <- 0.05/nrow(adjusted_meta_2)
quantile(adjusted_meta_2$p.value, probs=c(0.01, .05, .1, .2, .3, .4, .5, .6, .7, .8, .9, .95, .99), na.rm = T)
quantile(adjusted_meta_2$estimate, probs=c(0.01, .05, .1, .2, .3, .4, .5, .6, .7, .8, .9, .95, .99), na.rm = T)
sum(adjusted_meta_2$pvalue_bonferroni < 0.05)/nrow(adjusted_meta_2) 

adjusted_meta_2 |> group_by(sig_levels) |> count()
adjusted_meta_2 |> filter(sig_levels == 'BY<0.05') |> arrange(-p.value) |> head()

## qqplot
pval_qq <- data.frame(observed = sort(adjusted_meta_2$p.value), expected = (1:nrow(adjusted_meta_2))/nrow(adjusted_meta_2))
qq_p <- ggplot(pval_qq, aes(-log10(expected), -log10(observed)))
qq_p <- qq_p + geom_point()
##

p <- ggplot(pval_qq, aes(observed))
p <- p + geom_density() + theme_bw()
p <- p + geom_hline(yintercept = 1, color='blue')
p <- p + xlab("P-E association pvalue")

p <- ggplot(adjusted_meta_2, aes(p.value))
p <- p + geom_density() + theme_bw() + facet_grid(~ecategory)
p1 <- p + xlab("P-E association pvalue")


p <- ggplot(adjusted_meta_2, aes(p.value))
p <- p + geom_density() + theme_bw() + facet_grid(~pcategory)
p2 <- p + xlab("P-E association pvalue")

plot_grid(p1, p2, nrow=2, labels=c("A", "B"))

```

# Association Size vs. -log10(pvalue)

```{r average across e p}

e_summary <- adjusted_meta_2 |> group_by(evarname) |> arrange(pvalue_bonferroni) |>  
  summarize(mean_r2=mean(mean_adjusted_base_r2_diff, na.rm=T),  mean_estimate=mean(abs(estimate), na.rm=T), 
            median_pvalue=median(p.value, na.rm=T), n_sig=sum(pvalue_bonferroni < 0.05, na.rm=T), 
            n_tests=sum(!is.na(pvalue_bonferroni)),  median_i.squared=median(i.squared, na.rm=T),
            max_r2=first(mean_adjusted_base_r2_diff), max_pvarname=first(pvarname) , max_estimate=first(estimate), max_p.value=first(p.value)) |> mutate(n_sig_pct=n_sig/n_tests)


p_summary <- adjusted_meta_2 |> group_by(pvarname) |> arrange(pvalue_bonferroni) |> 
  summarize(mean_r2=mean(mean_adjusted_base_r2_diff, na.rm=T), mean_estimate=mean(abs(estimate), na.rm=T),
            median_pvalue=median(p.value, na.rm=T), n_sig=sum(pvalue_bonferroni < 0.05, na.rm=T), 
            n_tests=sum(!is.na(pvalue_bonferroni)),  median_i.squared=median(i.squared, na.rm=T),
            max_r2=first(mean_adjusted_base_r2_diff), max_evarname=first(evarname) , max_estimate=first(estimate), max_p.value=first(p.value)) |> mutate(n_sig_pct=n_sig/n_tests)

## deeper summary by group

p_group_summary <- adjusted_meta_2 |> unite(p_scategory, c(pcategory, psubcategory)) |> group_by(p_scategory) |> arrange(pvalue_bonferroni) |>  
  summarize(mean_r2=mean(mean_adjusted_base_r2_diff, na.rm=T),  mean_estimate=mean(abs(estimate), na.rm=T), 
            median_pvalue=median(p.value, na.rm=T), n_sig=sum(pvalue_bonferroni < 0.05, na.rm=T), 
            n_tests=sum(!is.na(pvalue_bonferroni)),  median_i.squared=median(i.squared, na.rm=T),
            max_r2=first(mean_adjusted_base_r2_diff), max_evarname=first(evarname) , max_estimate=first(estimate), max_p.value=first(p.value)) |> mutate(n_sig_pct=n_sig/n_tests)


e_group_summary <- adjusted_meta_2 |> unite(e_scategory, c(ecategory, esubcategory)) |> group_by(e_scategory) |> arrange(pvalue_bonferroni) |>  
  summarize(mean_r2=mean(mean_adjusted_base_r2_diff, na.rm=T),  
            mean_abs_estimate=mean(abs(estimate), na.rm=T),
            mean_estimate=mean((estimate), na.rm=T),
            median_pvalue=median(p.value, na.rm=T), 
            n_sig=sum(pvalue_bonferroni < 0.05, na.rm=T), 
            n_tests=sum(!is.na(pvalue_bonferroni)),  
            median_i.squared=median(i.squared, na.rm=T),
            max_r2=first(mean_adjusted_base_r2_diff), 
            max_pvarname=first(pvarname), 
            max_estimate=first(estimate), max_p.value=first(p.value)) |> mutate(n_sig_pct=n_sig/n_tests)





```


# Tables 1 and 2
```{r}
## 
e_group_summary <- adjusted_meta_2 |> filter(expo_name %in% c('expo', 'expo1', 'expo2', 'expo3')) |> filter(mean_adjusted_base_r2_diff <= .1) |> unite(e_scategory, c(ecategory, esubcategory)) |> group_by(e_scategory) |> arrange(pvalue_bonferroni,mean_adjusted_base_r2_diff) |>  
  summarize(
            #median_r2=mean(mean_adjusted_base_r2_diff, na.rm=T),  
            #median_abs_estimate=median(abs(estimate), na.rm=T),
            n_sig=sum(pvalue_bonferroni < 0.05, na.rm=T), 
            n_tests=sum(!is.na(pvalue_bonferroni)),  
            #median_i.squared=median(i.squared, na.rm=T),
            max_r2=first(mean_adjusted_base_r2_diff), 
            max_termname=remove_units_from_string(first(expo_name)),
            max_pvarname=remove_units_from_string(first(pvardesc)),
            max_evarname=remove_units_from_string(first(evardesc)),
            max_estimate=first(estimate), max_p.value=first(p.value), max_i.squared=first(i.squared)) |> mutate(n_sig_pct=n_sig/n_tests)

e_bonf_group_summary <- adjusted_meta_2 |> filter(sig_levels == 'Bonf.<0.05') |> 
  unite(e_scategory, c(ecategory, esubcategory)) |>
  mutate(sgn=ifelse(sign(estimate) <0, 'neg', 'pos')) |> group_by(e_scategory, sgn) |> arrange(pvalue_bonferroni) |>  
  summarize(median_bonf_r2=median(mean_adjusted_base_r2_diff, na.rm=T),  
            q25_bonf_estimate=quantile(estimate, probs=.25, na.rm=T),
            median_bonf_estimate=median((estimate), na.rm=T),
            q75_bonf_estimate=quantile(estimate, probs=.75, na.rm=T)
  )

e_bonf_group_summary <- e_bonf_group_summary |> pivot_wider(names_from=sgn, values_from=(c(median_bonf_r2, q25_bonf_estimate, median_bonf_estimate, q75_bonf_estimate)))

## merge
e_group_summary <- e_group_summary |> left_join(e_bonf_group_summary, by='e_scategory')
e_group_summary <- e_group_summary |> separate(col=e_scategory, into=c("escategory", "esubcategory"), sep="_")
e_group_summary <- e_group_summary |> filter(n_sig > 1)

e_group_summary <- e_group_summary |> select(escategory, esubcategory, n_tests, n_sig_pct, median_bonf_r2_neg, median_bonf_r2_pos, q25_bonf_estimate_neg, median_bonf_estimate_neg, q75_bonf_estimate_neg, q25_bonf_estimate_pos, median_bonf_estimate_pos, q75_bonf_estimate_pos, 
max_pvarname, max_evarname, max_estimate, max_r2, max_p.value, max_i.squared)

e_group_summary<- e_group_summary |> gt() |> 
  fmt_number(columns = c(starts_with("q"), starts_with("med")), decimals = 3) |>
  fmt_number(columns = c(max_r2, max_estimate), decimals = 3) |>
  fmt_number(columns = c(n_sig_pct), decimals = 2) |>
  fmt_number(columns = c(max_i.squared), decimals = 0) |>
  fmt_scientific(columns = max_p.value, decimals = 0)

e_group_summary <- e_group_summary |> 
  tab_spanner(label = "Effect Sizes < 0",columns = ends_with("neg")) |>
  tab_spanner(label = "Effect Sizes > 0",columns = ends_with("pos"))

e_group_summary <- e_group_summary |> 
  tab_spanner(label = "Example P-E Association (lowest p.value for E domain)",columns = starts_with("max"))

e_group_summary |> gtsave('./e_group_summary.html')
e_group_summary
```

```{r}
## 
p_group_summary <- adjusted_meta_2 |> filter(expo_name %in% c('expo', 'expo1', 'expo2', 'expo3')) |> filter(mean_adjusted_base_r2_diff <= .1) |> unite(p_scategory, c(pcategory, psubcategory)) |> group_by(p_scategory) |> arrange(pvalue_bonferroni,mean_adjusted_base_r2_diff) |>  
  summarize(
            #median_r2=mean(mean_adjusted_base_r2_diff, na.rm=T),  
            #median_abs_estimate=median(abs(estimate), na.rm=T),
            n_sig=sum(pvalue_bonferroni < 0.05, na.rm=T), 
            n_tests=sum(!is.na(pvalue_bonferroni)),  
            #median_i.squared=median(i.squared, na.rm=T),
            max_r2=first(mean_adjusted_base_r2_diff), 
            max_termname=remove_units_from_string(first(expo_name)),
            max_pvarname=remove_units_from_string(first(pvardesc)),
            max_evarname=remove_units_from_string(first(evardesc)),
            max_estimate=first(estimate), max_p.value=first(p.value), max_i.squared=first(i.squared)) |> mutate(n_sig_pct=n_sig/n_tests)

p_bonf_group_summary <- adjusted_meta_2 |> filter(sig_levels == 'Bonf.<0.05') |> 
  unite(p_scategory, c(pcategory, psubcategory)) |>
  mutate(sgn=ifelse(sign(estimate) <0, 'neg', 'pos')) |> group_by(p_scategory, sgn) |> arrange(pvalue_bonferroni) |>  
  summarize(median_bonf_r2=median(mean_adjusted_base_r2_diff, na.rm=T),  
            q25_bonf_estimate=quantile(estimate, probs=.25, na.rm=T),
            median_bonf_estimate=median((estimate), na.rm=T),
            q75_bonf_estimate=quantile(estimate, probs=.75, na.rm=T)
  )

p_bonf_group_summary <- p_bonf_group_summary |> pivot_wider(names_from=sgn, values_from=(c(median_bonf_r2, q25_bonf_estimate, median_bonf_estimate, q75_bonf_estimate)))

## merge
p_group_summary <- p_group_summary |> left_join(p_bonf_group_summary, by='p_scategory')
p_group_summary <- p_group_summary |> separate(col=p_scategory, into=c("pscategory", "psubcategory"), sep="_")
p_group_summary <- p_group_summary |> filter(n_sig > 1)

p_group_summary <- p_group_summary |> select(pscategory, psubcategory, n_tests, n_sig_pct, q25_bonf_estimate_neg, median_bonf_estimate_neg, q75_bonf_estimate_neg, q25_bonf_estimate_pos, median_bonf_estimate_pos, q75_bonf_estimate_pos, 
max_pvarname, max_evarname, max_estimate, max_r2, max_p.value, max_i.squared)

p_group_summary<- p_group_summary |> gt() |> 
  fmt_number(columns = c(starts_with("q"), starts_with("med")), decimals = 3) |>
  fmt_number(columns = c(max_r2, max_estimate), decimals = 3) |>
  fmt_number(columns = c(n_sig_pct), decimals = 2) |>
  fmt_number(columns = c(max_i.squared), decimals = 0) |>
  fmt_scientific(columns = max_p.value, decimals = 0)

p_group_summary <- p_group_summary |> 
  tab_spanner(label = "Effect Sizes < 0",columns = ends_with("neg")) |>
  tab_spanner(label = "Effect Sizes > 0",columns = ends_with("pos"))

p_group_summary <- p_group_summary |> 
  tab_spanner(label = "Example P-E Association (lowest p.value for E domain)",columns = starts_with("max"))

p_group_summary |> gtsave('./p_group_summary.html')
p_group_summary
```

```{r volcano, fig.height=8, fig.width=8}
adjusted_meta_2 <- adjusted_meta_2 |> mutate(p_cap = ifelse(p.value < 1e-30, 1e-30, p.value))
p <- ggplot(adjusted_meta_2 |> filter(ecategory != 'autoantibody'), aes(estimate, -log10(p_cap)))
p <- p + geom_point(shape='.') + scale_x_continuous(limits=c(-1, 1))
p <- p + facet_grid(ecategory ~ .) + scale_color_npg()
p <- p + geom_hline(yintercept = -log10(.05/nrow(adjusted_meta_2)), color='lightblue')
p <- p + theme_minimal() + theme(legend.position = "none") +ylab('p.value') + xlab("estimate")
p1 <- p

## uBiome only
p <- ggplot(adjusted_meta_2 |> filter(pcategory == 'microbiome'), aes(estimate, -log10(p_cap)))
p <- p + geom_point(shape='.') + scale_x_continuous(limits=c(-1, 1))
p <- p + facet_grid(ecategory ~ .) + scale_color_npg()
p <- p + geom_hline(yintercept = -log10(.05/nrow(adjusted_meta_2)), color='lightblue')
p <- p + theme_minimal() + theme(legend.position = "none") +ylab('p.value') + xlab("estimate")


p <- ggplot(adjusted_meta_2, aes(estimate, -log10(p_cap)))
p <- p + geom_point(shape='.') + scale_x_continuous(limits=c(-1, 1))
p <- p + facet_grid(pcategory ~ .) + scale_color_npg()
p <- p + geom_hline(yintercept = -log10(.05/nrow(adjusted_meta_2)), color='lightblue')
p <- p + theme_minimal() + theme(legend.position = "none") +ylab('p.value') + xlab("estimate")
p2 <-p

plot_grid(p1, p2, ncol=2, labels=c("A", "B"))
```


```{r}
e_effect_sizes_per <- adjusted_meta_2 |> filter(sig_levels == 'Bonf.<0.05') |> group_by(ecategory, esubcategory, sign(estimate)) |> summarize(median_pvalue=median(p.value), number_signficant=n(), mean_estimate=mean((estimate))) |> arrange(-mean_estimate)

e_effect_sizes_per <- e_effect_sizes_per |> mutate(esubcategory = ifelse(is.na(esubcategory), ecategory, esubcategory))
p <- ggplot(e_effect_sizes_per, aes(mean_estimate, -log10(median_pvalue), label=esubcategory))
p <- p + geom_point(aes(size=number_signficant)) + geom_text_repel() + geom_vline(xintercept = 0)
p <- p + theme_bw() + xlab("Average(Estimate) within exposome groups") + ylab("Median log10(pvalue)")
p <- p + theme(legend.position = "bottom")
p

p_effect_sizes_per <- adjusted_meta_2 |> filter(sig_levels == 'Bonf.<0.05') |> group_by(pcategory, psubcategory) |> summarize(mean_r2 = mean(mean_adjusted_base_r2_diff, na.rm=T))

```

```{r cdfs}

p <- ggplot(adjusted_meta_2 |> filter(vartype =='categorical'), aes(abs(estimate), color=sig_levels))
p <- p + stat_ecdf() + scale_x_continuous(limits=c(0, .25))
p <- p + xlab("abs(estimate)") + ylab("percentile")  + theme(legend.position="bottom") + scale_color_npg()
p

p <- ggplot(adjusted_meta_2 |> filter(vartype =='continuous'), aes(abs(estimate), color=sig_levels))
p <- p + stat_ecdf() + scale_x_continuous(limits=c(0, .25))
p <- p + xlab("abs(estimate)") + ylab("percentile") + theme(legend.position="bottom") + scale_color_npg()
p


p <- ggplot(adjusted_meta_2, aes(abs(estimate), color=sig_levels))
p <- p + stat_ecdf() + scale_x_continuous(limits=c(0, .25))
p <- p + xlab("abs(estimate)") + ylab("percentile") + theme(legend.position="bottom") + scale_color_npg()
p


ecdf_for_sig <- adjusted_meta_2 |> filter(sig_levels == 'Bonf.<0.05') |> pull(mean_adjusted_base_r2_diff) |> ecdf()
ecdf_for_non_sig <- adjusted_meta_2 |> filter(sig_levels == '> BY & Bonf.') |> pull(mean_adjusted_base_r2_diff) |> ecdf()


p_effect_sizes_per <- p_effect_sizes_per |> mutate(q = ecdf_for_sig(mean_r2), sig_levels ='Bonf.<0.05')
p_effect_sizes_per <- p_effect_sizes_per |> mutate(p_cat = ifelse(is.na(psubcategory), pcategory, psubcategory))
p <- ggplot(adjusted_meta_2, aes(mean_adjusted_base_r2_diff, color=sig_levels))
p <- p + stat_ecdf() + scale_x_continuous(limits=c(0, .05)) +scale_color_aaas()
p <- p + geom_point(data=p_effect_sizes_per, aes(x=mean_r2, y = q, color=sig_levels)) 
p <- p + geom_text_repel(data=p_effect_sizes_per, aes(x=mean_r2, y = q, color=sig_levels, label=p_cat)) 
p <- p + xlab("R^2 (adjusted-base model)") + ylab("percentile") 
p <- p + theme_bw() + theme(legend.position="bottom") 
p


```

# Association sizes for all exposures

## contextualizing all exposures

```{r}
adjusted_meta_2 <- adjusted_meta_2 |> mutate(evarname = fct_reorder(evarname, abs(estimate), mean))
#p <- ggplot(adjusted_meta_2, aes(x=evarname, y=estimate, group=evarname))
#p <- p + geom_density_ridges()
#p <- p + geom_boxplot() + facet_grid(~sig_levels)
#p
```

# Replicability and Consistency

## I2 across number of surveys for P-E pair

```{r i2}
## reorder sig_levels
adjusted_meta_2 <- adjusted_meta_2 |> mutate(sig_levels=fct_relevel(sig_levels, c("> BY & Bonf.","BY<0.05", "Bonf.<0.05")))

p <- ggplot(adjusted_meta_2, aes(factor(nobs), i.squared,color=sig_levels))
p <- p + geom_boxplot() + xlab("Number of Surveys for PE Association") + theme(legend.position="bottom") + scale_color_aaas()
p <- p + theme_bw() + theme(legend.position="bottom") + ylab("i-squared")
p

i2_medians <- adjusted_meta_2 |> group_by(sig_levels) |> summarize(i2_median=median(i.squared))

```

# Reverse meta-analysis: replicability of P-E

```{r}
within_survey_pvalue_threshold <- 0.05 # should this be changed for each survey?
p_val_pe_pair <- expos_wide |> group_by(evarname, pvarname, term) |> summarize(n_pvalue_lt=sum(p.value_adjusted<within_survey_pvalue_threshold), total_n=n(), pct_pvalue_lt=n_pvalue_lt/total_n)
adjusted_meta_3 <- adjusted_meta_2 |> left_join(p_val_pe_pair, by=c("evarname"="evarname", "pvarname"="pvarname", "expo_name"="term"))
```

# Showcasing associations:

-   Low pvalue, Higher R2, low I2
-   Low pvalue, Higher R2, and higher I2

```{r}

adjusted_meta_3 |> filter(pvalue_bonferroni < 0.05) |> nrow() 
adjusted_meta_3 |> filter(pvalue_bonferroni < 0.05, n_pvalue_lt >= 2) |> nrow()
non_het_pairs <- adjusted_meta_3 |> filter(pvalue_bonferroni < 0.05, n_pvalue_lt >= 2, i.squared < 50, mean_adjusted_base_r2_diff > .025)
het_pairs <- adjusted_meta_3 |> filter(pvalue_bonferroni < 0.05, n_pvalue_lt >= 2, i.squared > 50, mean_adjusted_base_r2_diff > .025, nobs >= 4)
#het_pairs_2 <- adjusted_meta_3 |> filter(sig_levels == 'BY<0.05', i.squared > 90, nobs >= 4) 

adjusted_meta_3 |> filter(pvalue_bonferroni < 0.05) |> group_by(ecategory) |> count()

## non-heterogeneous example

plot_pair <- function(evarname_str, pvarname_str, estimate_limits=c(0.01,.35)) {
  test_1 <- expos_wide |> filter(evarname == evarname_str, pvarname == pvarname_str) |> select(Begin.Year,  exposure_table_name, phenotype_table_name, e_variable_description, p_variable_description, estimate_adjusted, std.error_adjusted, p.value_adjusted) 
  exposure_name <- remove_units_from_string(test_1$e_variable_description[1])
  phenotype_name <- remove_units_from_string(test_1$p_variable_description[1])
  test_1 <- test_1 |> select(Begin.Year, estimate_adjusted, std.error_adjusted) |> rename(estimate=estimate_adjusted, std.error = std.error_adjusted, Survey=Begin.Year) |> mutate(i.squared = NA, i.squared_text='')
meta_test_1 <- adjusted_meta_2 |> filter(evarname == evarname_str, pvarname == pvarname_str) |> mutate(Survey = 'overall') |> select(Survey, estimate, std.error, i.squared) |> mutate(i.squared_text = sprintf("%i%%", round(i.squared)))
  test_1 <- test_1 |> rbind(meta_test_1) 
  test_1 <- test_1 |> mutate(point_shape = ifelse(Survey == 'overall', 23, 21)) |> mutate(point_shape = as.integer(point_shape)) 
  test_1 <- test_1 |> mutate(point_size = ifelse(Survey == 'overall', 7, 2)) 
  p <- ggplot(test_1, aes(Survey, estimate))
  p <- p + geom_point(aes(shape=point_shape, size=point_size, fill=point_shape)) + scale_shape_identity() + scale_size_identity()
  p <- p + geom_text(data=meta_test_1, aes(Survey, estimate, label=i.squared_text), size=3, color="black", nudge_x=.6) 
  p <- p + geom_errorbar(aes(ymin=estimate-1.96*std.error, ymax=estimate+1.96*std.error), width=.1) + scale_x_discrete(limits=rev)
  p <- p + scale_y_continuous(limits=estimate_limits)
  p <- p + coord_flip()  + theme_bw() + theme(legend.position = "none")
  p <- p + ggtitle(sprintf('scale(%s)-scale(log10(%s))', phenotype_name, exposure_name))+ theme(plot.title = element_text(size = 7))
}

## non-heterogeneous example
p1 <- plot_pair('URXP01', 'LBDNENO')
## heterogeneous example
p2 <- plot_pair('LBXGTC', 'BMXBMI')

p3 <- plot_pair('LBXBPB', 'BMXHT', c()) # 33% i2

p4 <- plot_pair('LBXCOT', 'BPXPLS', c()) # 33% i2



#expos_wide |> filter(evarname == 'LBXPFOS', pvarname == 'LBXSAL') 
plot_grid(p1, p2, p3, p4, ncol=2,labels = c('A', 'B', "C", "D"), label_size = 12)

```

 # Examples for paper of top hits
```{r}
p1 <- plot_pair('LBXPFHS', 'LBXSAL') + ylab("Association Estimate")
## heterogeneous example
p2 <- plot_pair('LBXGTC', 'LBXTC') + ylab("Association Estimate")

p3 <- plot_pair('LBXBPB', 'BMXWT', c(-.3, 0)) + ylab("Association Estimate")

p4 <- plot_pair('LBXCOT', 'LBDNENO')+ ylab("Association Estimate")

plot_grid(p1, p2, p3, p4, ncol=2,labels = c('A', 'B', "C", "D"), label_size = 12)

```

```{r}
library(ggridges)
## show histogram of associations for all top findings for  LBXGTC, "LBXBPB", "LBXPFHS", "LBXCOT"
ep_candidates <-  tibble(evarname = c("LBXGTC", "LBXBPB", "LBXPFHS", "LBXCOT"), 
                         pvarname = c("LBXTC", "BMXWT", "LBXSAL", "LBDNENO"))
exposure_dist <- adjusted_meta_2 |> filter(evarname %in% ep_candidates$evarname) |> filter(sig_levels == "Bonf.<0.05") |> mutate(evardesc = remove_units_from_string(evardesc))

## collect Survey specific associations
survey_exposure_pts <- vector(mode = "list", length=nrow(ep_candidates))
for(rw_num in 1:nrow(ep_candidates)) {
  survey_exposure_pts[[rw_num]] <- expos_wide |> filter(evarname == ep_candidates$evarname[rw_num], pvarname == ep_candidates$pvarname[rw_num]) |> 
  select(Begin.Year,  evarname, pvarname, exposure_table_name, phenotype_table_name, e_variable_description, p_variable_description, estimate_adjusted, std.error_adjusted, p.value_adjusted) 
}

survey_exposure_pts <- survey_exposure_pts |> bind_rows() |> mutate(evardesc=remove_units_from_string(e_variable_description))

exposure_dist <- exposure_dist |> mutate(evarname = substr(evarname, 4, 8)) 
survey_exposure_pts <- survey_exposure_pts |> mutate(pvarname = substr(pvarname, 4, 8), evarname = substr(evarname, 4, 8))

p <- ggplot(exposure_dist, aes(y=evarname, x=estimate, fill=evarname))
p <- p + geom_density_ridges()+ scale_fill_jama(guide="none") #+ scale_colour_continuous(guide = "none")
p <- p + geom_point(data=survey_exposure_pts, aes(y=evarname, x=estimate_adjusted, color=factor(Begin.Year))) + scale_color_aaas(name = "")
pheno_het <- p + theme_bw()  + theme(legend.position='bottom') + ylab("Exposure Name") + xlab("Association Estimate")

survey_het <- plot_grid(p1, p2, p3, p4, ncol=2,labels = c('A', 'B', "C", "D"), label_size = 12)

p <- plot_grid(survey_het, pheno_het, ncol=2, labels=c("", "E"), label_size = 12)

```

# R2 adjusted vs. non-adjusted

```{r}

p <- ggplot(expos_wide, aes(rsq_base, rsq_adjusted))
p <- p + geom_point(shape='.') + xlab("R2 Base [Demographics]") + ylab("R2 [Exposure + Demographics]")
p <- p + theme_bw()
p

p <- ggplot(expos_wide, aes(rsq_adjusted, rsq_unadjusted))
p <- p + geom_point(shape='.') + xlab("R2 Adjusted [Exposure + Demographics]") + ylab("R2 Unadjusted [Exposure]")
p <- p + theme_bw() 
p

p <- ggplot(adjusted_meta_2, aes(mean_adjusted_base_r2_diff, i.squared))
p <- p + geom_point(shape='.') + xlab("R2 Adjusted [Exposure + Demographics]") + ylab("i.squared") + scale_x_continuous(limits=c(0, .01))
p <- p + theme_bw() + facet_grid(~sig_levels)
p

```

# Estimate and P-values: adjusted vs. non-adjusted

-   estimate the association globally between adjusted and non-adjusted models

```{r fig.height=6, fig.width=9}
p <- ggplot(expos_wide, aes(estimate_unadjusted, estimate_adjusted))
p <- p + geom_point(shape='.') + scale_x_continuous(limits=c(-2, 2)) + scale_y_continuous(limits=c(-2, 2)) + xlab("Adjusted [Exposure + Demographics]") + ylab("Unadjusted [Exposure]") + geom_abline()
p <- p + geom_smooth(method="lm")
p <- p + theme_bw()
p

## how much bias?
tidy(lm(estimate_adjusted  ~ estimate_unadjusted, expos_wide)) # biased to be lower (intercept is negative, and slope is less than 1)
# 


p <- ggplot(expos_wide, aes(-log10(p.value_unadjusted), -log10(p.value_adjusted)))
p <- p + geom_point(shape='.', alpha=.1) + xlab("Adjusted model pvalue [Exposure + Demographics]") + ylab("Unadjusted Pvalue [Exposure]")
p <- p + geom_abline()
p <- p + theme_bw()
p


p <- ggplot(expos_wide, aes(statistic_unadjusted, statistic_adjusted))
p <- p + geom_point(shape='.', alpha=.1) + xlab("Adjusted model z [Exposure + Demographics]") + ylab("Unadjusted z [Exposure]")
p <- p + geom_abline() + scale_y_continuous(limit=c(-20, 20)) + scale_x_continuous(limits=c(-20,20))
p <- p + theme_bw()
p

p <- ggplot(adjusted_meta_2 , aes(estimate, estimate_unadjusted, color=sig_levels))
p <- p + geom_point(shape='.') + scale_x_continuous(limits=c(-1, 1)) + scale_y_continuous(limits=c(-1, 1)) + scale_color_aaas()
p <- p + geom_abline()
p <- p + facet_grid(~sig_levels) + xlab("Adjusted model estimate [Exposure + Demographics]") + ylab("Unadjusted estimate [Exposure]")
p <- p + geom_smooth(method="lm")
p <- p + theme_bw() +theme(legend.position = "none")
p



tidy(lm(estimate ~ estimate_unadjusted, adjusted_meta_2))
tidy(lm(estimate ~ estimate_unadjusted, adjusted_meta_2 |> filter(pvalue_bonferroni < .05)))
tidy(lm(estimate ~ estimate_unadjusted, adjusted_meta_2 |> filter(sig_levels == '> BY & Bonf.')))
tidy(lm(estimate ~ estimate_unadjusted, adjusted_meta_2 |> filter(sig_levels == 'BY<0.05')))


p <- ggplot(adjusted_meta_2 , aes(estimate_unadjusted, estimate, color=ecategory))
p <- p + geom_point(shape='.') + scale_x_continuous(limits=c(-1, 1)) + scale_y_continuous(limits=c(-1, 1)) + scale_color_aaas()
p <- p + geom_abline()
p <- p + facet_grid(~ecategory) + ylab("Adjusted model estimate [Exposure + Demographics]") + xlab("Unadjusted estimate [Exposure]")
p <- p + geom_smooth(method="lm")
p <- p + theme_bw() +theme(legend.position = "none")
p


```

# Which e and P domains are most subject to demographic bias?

-   nutrients and pollutants

```{r}
bias_per_ecategory <- adjusted_meta_2 |> group_by(ecategory) |> summarize(
  mod=list(lm(estimate ~ estimate_unadjusted))) |> mutate(tidied=map(mod, tidy)) |> unnest(tidied)

bias_per_ecategory |> select(-mod) |> select(ecategory, term, estimate, p.value) |> filter(term == "estimate_unadjusted") |> select(-term)


bias_per_pcategory <- adjusted_meta_2 |> group_by(pcategory) |> summarize(
  mod=list(lm(estimate ~ estimate_unadjusted))) |> mutate(tidied=map(mod, tidy)) |> unnest(tidied)

bias_per_pcategory |> select(-mod) |> select(pcategory, term, estimate, p.value) |> filter(term == "estimate_unadjusted") |> select(-term)

```

# Multivariate R2 of the exposome

```{r r2}
p <- ggplot(mvr2, aes(n_evars, mv))
p <- p + geom_point() + theme_bw()
p <- p + geom_text(data=mvr2 |> filter(n_evars > 40),aes(n_evars, mv, label=pvarname) )
p <- p + xlab("Number of Exposome Variables in Model") + ylab("R-squared")
p

mvr2 |> collect() |> summarize(mean_sample_size=mean(n), median_r2=median(mv), q25_r2=quantile(mv, probs=.25), q75_r2=quantile(mv, probs=.75))

```

# Correlation of phenotypes in exposome space

```{r}
library(corrr)
library(gplots)
to_array <- adjusted_meta_2 |> filter(expo_name == 'expo', vartype == 'continuous') |> select(evarname, pvarname, estimate, p.value) |> mutate(estimate= ifelse(p.value >= 0.05, 0, estimate)) |> mutate(estimate = ifelse(is.na(estimate), 0, estimate))  |> select(-p.value) |> pivot_wider(names_from = pvarname, values_from = estimate)

exposure_correlation <- to_array |> select(-evarname) |> correlate(diagonal = 1)
m <- exposure_correlation |> as_matrix()
m[is.na(m)] <- 0

heatmapColors <- function(numColors=16) {
	c1 <- rainbow(numColors,v=seq(0.5,1,length=numColors),s=seq(1,0.3,length=numColors),start=4/6,end=4.0001/6);
	c2 <- rainbow(numColors,v=seq(0.5,1,length=numColors),s=seq(1,0.3,length=numColors),start=1/6,end=1.0001/6);
	c3 <- c(c1,rev(c2)); 
	return(c3)
}

heatmap.2(m, trace = 'none', na.rm = F, scale = 'none', symm = T, col=heatmapColors(5), margins=c(16,16), sepwidth=c(.1, .1), symbreaks=T)

```

# Correlation of the exposome

```{r}

to_array <- adjusted_meta_2 |> filter(expo_name == 'expo', vartype == 'continuous') |> select(evarname, pvarname, estimate, p.value) |> mutate(estimate= ifelse(p.value >= 0.05, 0, estimate)) |> mutate(estimate = ifelse(is.na(estimate), 0, estimate))  |> select(-p.value) |> pivot_wider(names_from = evarname, values_from = estimate)

phenome_correlation <- to_array |> select(-pvarname) |> correlate(diagonal = 1)
m <- phenome_correlation |> as_matrix()
m[is.na(m)] <- 0

heatmap.2(m, trace = 'none', na.rm = F, scale = 'none', symm = T, col=heatmapColors(5), margins=c(16,16), sepwidth=c(.1, .1), symbreaks=T)


```
