
---
title: "High level tour: Running an NHANES Exposure-Wide Association Study (ExWAS)"
author: "nhanespewas team"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_depth: 2
---

> **Goal**  
> Show, in a few reproducible steps, how the **exwas.R** driver script in the  
> [`nhanespewas`](https://github.com/chiragjp/nhanespewas) package performs an  
> Exposome-Wide scan for one phenotype.

---

## 1  What the script does

`exwas.R` loops over every eligible exposure for a chosen phenotype and:

1. **Filters** pairs to those with ≥ 500 participants in ≥ 2 NHANES cycles.  
2. **Chooses** the right transformation for each exposure  
   (log-scale, quantile bins, or categorical).  
3. **Builds** nine predefined adjustment models (age-only, fully adjusted, etc.).  
4. **Fits** survey-weighted regressions with `pe_flex_adjust()` and captures  
   tidy coefficients, model “glance” statistics, and incremental R².  
5. **Meta-analyses** results across cycles with unrestricted weighted least squares  
   (UWLS), adding heterogeneity metrics (Q, I², H²).  
6. **Saves** one compressed **RDS** file containing the three result tables  
   (`pe_tidied`, `pe_glanced`, `rsq`) plus raw model objects for audit.

---

## 2  Setup

```{r eval=FALSE, include=TRUE}
# install.packages(c("remotes", "DBI", "RSQLite", "logger",
#                    "getopt", "tidyverse"))
remotes::install_github("chiragjp/nhanespewas")

library(nhanespewas)
```

You also need
	- nhanes_XXXX.sqlite – a SQLite database built by the package’s helper, available here: https://doi.org/10.6084/m9.figshare.29182196.v1 (~2GB)
	-	sample_size_pe_category_0824.csv – shipped in the package’s extdata folder.

⸻

## 3  A minimal command-line call
```{bash eval=FALSE, include=TRUE}
Rscript exwas.R -p LBXRDW -i ../db/nhanes_031725.sqlite -o .
```

This scans all 619 exposures for the phenotype LBXRDW and writes
./LBXRDW.rds.

Restrict to selected exposures + quantile modelling
 exposures_to_run.csv: one column, e.g. LBXBCD, URXUHG


⸻


## 4  Exploring the output


```{r eval=FALSE, include=TRUE}
res <- readRDS("LBXRDW.rds")
names(res)
head(res$pe_tidied[, c("evarname","estimate","p.value","pval_BY")])
```

res element	description

- pe_tidied	coefficient, SE, p value for each model/level
- pe_glanced	model-wise fit stats (AIC, df, etc.) for each model
- rsq	incremental R² for each adjustment scenario
- modls	raw model lists (one per exposure)

⸻

## 5  Algorithm flow

```{r echo=FALSE}
knitr::kable(
  tibble::tribble(
    ~Step, ~Function, ~Output,
    "1 Filter pairs",    "dplyr::filter()",    "≥500 N, ≥2 cycles",
    "2 Detect type",     "check_e_data_type()", "continuous / categorical / rank",
    "3 Build models",    "adjustment_scenario_for_variable()", "9 covariate sets",
    "4 Fit survey GLM",  "pe_flex_adjust()",  "tidy / glance / R²",
    "5 Meta-analyse",    "internal UWLS",     "β_UWLS, p_UWLS, I²",
    "6 Save results",    "saveRDS()",         "<phenotype>[ _subset].rds"
  ),
  caption = "High-level workflow executed by **exwas.R**"
)
```



