---
title: "Exploring the Database and Catalogs"
author: "Chirag Patel"
date: "2025-06-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this .rmd, we will show where the meta-data are located, e.g,, how to find what variables are in what tables in the NHANES database


```{r}
library(tidyverse) # for data‐manipulation (dplyr , etc.)
library(survey) # loaded by nhanespewas for complex‐survey regressions
library(nhanespewas)
library(DBI)
```

# Explore the NHANES database

- each table name corresponds to a type of measurement
- the NHANES are  grouped into different survey years, and the survey sequence is appended to the table, e.g.:
- A: 1999-2000, B: 2001-2002, C: 2003-2004, D: 2005-2006, E: 2007-2008, F: 2009-2010, G: 2011-2012, H: 2013-2014, I: 2014-2015, J: 2016-2017
- each survey uear contains an independent snapshot of the US population.
- the "demo_X" tables contain the both the entirety of total participants of a certain survey year along with demographic attributes (age, sex, education, income)

```{r}
# Replace './db/nhanes_031725.sqlite' with your actual file path
path_to_db <- "./db/nhanes_031725.sqlite"
# Establish a connection to the NHANES PE database
con <- connect_pewas_data(path_to_db)
# Once connected, you can reference tables via dplyr::tbl()
varnames <- dplyr::tbl(con, "variable_names_epcf") 

table_names <- tibble(table_name=dbListTables(con))
table_names # see column Variable.Name and Data.File.Name

varnames # this table is an index of variables in tables
# demographic table for 2001-2002:
demo_b <- tbl(con, "demo_b")
```

# Built-in Exposome and Phenome Catalogs

- we have compiled a list of exposures, their descriptions, and the categories they belong to
- we have done the analogous for phenotypes

```{r explore the catalog}
data("p_catalog") # load them into the memory
data("e_catalog")

e_catalog
p_catalog

```

# "Select" file: a list of exposure-phenotype pairs

- Now that you have seen a listing of phenotypes and exposures, which ones can you correlate?
- In the following file, we have a list of sample sizes by phenotype and exposure pairs and the tables they belong to (e_table_name, p_table_name)
- This file is used as input to pipeline code (see: pipeline/exwas.R) to run associations
- the pvarname is the phenotype NHANES name; the evarname is the NHANES exposure name
- n is the sample size for the pair of variables (the possible total N for an association)

```{r read in select pairs file}
select_filename <- system.file("extdata", "sample_size_pe_category_0824.csv", package = "nhanespewas")
select <- read_csv(select_filename)
select
```
# Run a sample association from the select file
```{r}

select_pair_1 <- select |> slice_head(n=1)
evarname <- select_pair_1 |> pull(evarname)
pvarname <- select_pair_1 |> pull(pvarname)
evarname
pvarname
test <- pe_flex_adjust(pvarname,evarname, adjustment_variables   = adjustment_models, con = con, scale_p = TRUE, logxform_p = TRUE )

test$models[[2]]$tidied # expo corresponds to the evarname exposure 

```

```{r}
dbDisconnect(con)
```
# Next steps

With the database and data structure described above, along with the "Select" file, you are now ready to start associaing each exposure to a phenotype. A script to do this for a phenotype is described in the code: pipeline/exwas.R


