---
title: "Quickstart: Associating a Phenotype with an Exposure"
author: "Chirag Patel"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# Set global chunk options: show code by default
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This vignette shows how to run a simple exposure–phenotype association
using the nhanespewas package. We will: 1. Load required libraries. 2.
Connect to a locally stored NHANES “PE” database (1999–2017). 3. Peek at
the predefined adjustment‐model objects. 4. Run three illustrative
associations (CRP vs. Cotinine, Hemoglobin A1c vs. gamma-tocopherol,
Lead vs. RDW). 5. Demonstrate how to include custom adjustment
scenarios. 6. Show one example of a binary‐exposure model (urinary
Chlamydia test positive vs. negative). 7. Finally, disconnect the
database.

Throughout, we illustrate how to retrieve tidy summary statistics, R²
values, and how to tweak scaling or log‐transformation options.



# 1.  Load Libraries

First, load the core packages you will need. • tidyverse: for
data‐manipulation functions (e.g., filter, tbl). • DBI: to manage the
database connection. • survey: (pulled in by nhanespewas internally for
survey‐based regressions). • nhanespewas: our package of interest.

```{r load}
library(tidyverse) # for data‐manipulation (dplyr, %\>%, etc.)
library(DBI) # for connecting to an SQLite (or other) database
library(survey) # loaded by nhanespewas for complex‐survey regressions
library(nhanespewas) # our package for PE associations
```


# 2.  Connect to the NHANES PE Database

We assume that you have already downloaded the SQLite database
containing all variables (1999–2017) formatted for nhanespewas. Adjust
the path_to_db to match your local file.

```{r connect}
# Replace './db/nhanes_031725.sqlite' with your actual file path

path_to_db <- "./db/nhanes_031725.sqlite"

# Establish a connection to the NHANES PE database

con <- connect_pewas_data(path_to_db)

# Once connected, you can reference tables via dplyr::tbl()

varnames <- dplyr::tbl(con, "variable_names_epcf")
```

What is variable_names_epcf? This table contains metadata for all NHANES
variables used in PE analyses: • Variable.Name (e.g., "LBXCRP") • Label
(e.g., “C-reactive protein (mg/dL)”) • Unit, category (Lab,
Questionnaire, etc.), and which survey cycles the variable appears in.



# 3.  Explore Predefined Adjustment Models

The nhanespewas package ships with several “adjustment‐model” templates
(e.g., age + sex + race/ethnicity, the fully adjusted model, etc.).
These are stored as a data frame named adjustment_models.

## Display the built‐in adjustment scenarios
```{r}
adjustment_models
```

What you’ll see: 
• A tibble with one row per scenario (e.g.,
age_sex_ethnicity, age_sex_ethnicity_income_education, etc.). 
• A variables column showing the R code (as a string) that identifies which
covariates to include. 
• A domain column (usually “default”).

Later, when you call pe_flex_adjust(), you can pass one of these
scenarios (or build your own) to specify exactly which covariates to
include.

# 4.  Example 1: C-Reactive Protein (P) vs. Cotinine (E)

Let’s run a simple linear regression of C-reactive protein (LBXCRP) on
serum cotinine (LBXCOT), adjusting for the default covariates and
scaling the phenotype. 1. Confirm that both variables exist in
variable_names_epcf. 2. Call pe_flex_adjust() with "LBXCRP" (phenotype),
"LBXCOT" (exposure), plus the adjustment template. 3. Extract summary
statistics and R² values.

## 1. Verify that LBXCOT and LBXCRP are present in the database

```{r}
varnames |> filter(Variable.Name == "LBXCOT") 
varnames |> filter(Variable.Name == "LBXCRP")
```

## 2. Run PE association: CRP (outcome) \~ Cotinine (exposure)

 - adjustment_models: built‐in set of covariates (age, sex, race_ethnicity, etc.)

 - scale_p = TRUE: z‐score the phenotype (LBXCRP) before regression

 - logxform_p = FALSE: do NOT log‐transform the phenotype

 - scale_type = 3: use RankNorm (type‐3 scaling)

```{r}
crp_cot <- pe_flex_adjust("LBXCRP","LBXCOT", adjustment_variables   = adjustment_models, con = con, scale_p = TRUE, logxform_p = FALSE, scale_type = 3)
fev1_cot <- pe_flex_adjust("SPXNFEV1","LBXCOT", adjustment_variables   = adjustment_models, con = con, scale_p = TRUE, logxform_p = FALSE, scale_type = 1)

```


## 3. Extract and print results

The `models` list holds two entries per survey cycle: base model (covariates only) and full model (covariates + exposure). Here, [[2]] corresponds to the fully adjusted model for the second NHANES wave.

```{r}
crp_cot$models[[2]]$tidied # coefficient, SE, p‐value, etc.
crp_cot$models[[2]]$r2 # R² object for the fully adjusted model
```

## 4. Compare against base model R²

```{r}
base_r2 <- crp_cot$base_models[[2]]$r2$rsq
full_r2  <- crp_cot$models[[2]]$r2$rsq 
delta_r2 <- full_r2 - base_r2
```

### Print out ΔR²
```{r}
delta_r2
```
Explanation of R² output: 
• base_models[[2]]$r2$rsq is the R² from the base model (covariates only) in cycle #2. 
• models[[2]]$r2$rsq is the R²
from the fully adjusted model (covariates + Cotinine). 
• ΔR² = full_r2 − base_r2 shows how much additional variance in CRP is explained by adding
cotinine.


# 5.  Example 2: Hemoglobin A1c (P) vs. Gamma-Tocopherol (E)

Next, we’ll examine the relationship between hemoglobin A1c (LBXGH) and
blood gamma-tocopherol (LBXGTC). Again, we use the same default adjustment set, plus SD scaling—no log‐transform.

# Run PE association: Hemoglobin A1c (LBXGH) \~ Gamma‐tocopherol (LBXGTC)

```{r}
gh_gtc <- pe_flex_adjust( "LBXGH", "LBXGTC", adjustment_variables = adjustment_models, con = con, scale_p = TRUE,logxform_p = FALSE )
```

# Display summary statistics for the fully adjusted model (model #2)

```{r}
gh_gtc$models[[2]]$tidied
```
# Extract R² for full vs. base model

```{r}
base_r2_gh <- gh_gtc$base_models[[2]]$r2$rsq
full_r2_gh  <- gh_gtc$models[[2]]$r2$rsq 
delta_r2_gh <- full_r2_gh -base_r2_gh
delta_r2_gh
```


Note: Once you run gh_gtc, the tibble gh_gtc$models[[2]]$tidied will
show the estimated β‐coefficient, standard error, p‐value, and sample
size for cycle #2 of NHANES.

# 6.  Example 3: Blood Lead (E) vs. RDW (P) with a Custom Adjustment Scenario

Sometimes you want to tweak the default adjustment set—for example, to
add survey cycle as a categorical factor or to include body‐mass index.
In this example, we’ll create a custom scenario named
"age_sex_ethnicity_income_education_plus_cycle", then run
pe_flex_adjust() for RDW (LBXRDW) \~ Blood lead (LBXBPB). 

## 1. Start by filtering down to the default scenario "age_sex_ethnicity_income_education". 

## 2. Append two additional covariates: 
• as.factor(SDDSRVYR): indicator for NHANES cycle 
• BMXBMI_adj: BMI (adjusted; already present in your code base) 

## 3. Run pe_flex_adjust() with scale_p = TRUE and also scale & log‐transform the exposure (LBXBPB).

### 1. Extract default scenario rows

```{r}
test_scenario <- adjustment_models |> filter(scenario == "age_sex_ethnicity_income_education")
```
### 2. Append two new variables to the same scenario:

- NHANES cycle indicator (as.factor(SDDSRVYR))

- Adjusted BMI (BMXBMI_adj)

```{r}
test_scenario <- test_scenario |> rbind( tibble(scenario ="age_sex_ethnicity_income_education", variables = "as.factor(SDDSRVYR)",domain = "default"), tibble(scenario = "age_sex_ethnicity_income_education", variables ="BMXBMI_adj", domain = "default") )
```

### 3. Run PE association: RDW (LBXRDW) \~ Blood lead (LBXBPB)

 - test_scenario: includes age, sex, race/ethnicity, income, education, cycle, BMI

 - scale_p = TRUE: z‐score RDW

 - scale_e = TRUE: z‐score blood lead

 - logxform_e = TRUE: log‐transform blood lead first

```{r}
rdw_bpb <- pe_flex_adjust("LBXRDW",  "LBXBPB", adjustment_variables =  test_scenario, con = con, scale_p = TRUE, scale_e = TRUE, logxform_e = TRUE )
```
Display summary statistics for the fully adjusted model 

```{r}
rdw_bpb$models[[1]]$tidied
```

 Compute ΔR² 

```{r}
base_r2_rdw <- rdw_bpb$base_models[[1]]$r2$rsq
full_r2_rdw  <- rdw_bpb$models[[1]]$r2$rsq 
delta_r2_rdw <- full_r2_rdw - base_r2_rdw
delta_r2_rdw
```


Why log‐transform and scale the exposure? 
• Many environmental biomarkers (like lead) have a skewed distribution. 
• Applying logxform_e = TRUE helps ease interpetation and linear modelings. 
• Afterwards, scale_e = TRUE standardizes the log‐transformed exposure so that coefficients are per-SD.

# 7.  Example 4: Binary Exposure (Urinary Chlamydia Test) and Urinary
    Creatinine (Continuous P)

In this example, we treat urinary Chlamydia test result (URXUCL) as a two‐level categorical exposure: “1” (negative) or “2” (positive). We want to see how urinary creatinine (URXUCR) differs between those two groups, adjusting for the default covariates.

## 1. Verify the Chlamydia variable exists and in what surveys

```{r}
varnames |> filter(Variable.Name == "URXUCL")
```
## 2. Run PE association: Urinary creatinine (URXUCR) \~ Urinary Chlamydia (URXUCL)

 - exposure_levels = c(1, 2): explicitly tell the function that “1”=negative, “2”=positive

 - scale_p = TRUE: z‐score the outcome (URXUCR)

 - scale_e = FALSE, logxform_e = FALSE: treat URXUCL as a two‐level categorical variable (no scaling or log‐transform)

```{r}
ucr_ucl <- pe_flex_adjust("URXUCR", "URXUCL", adjustment_variables =adjustment_models, con = con, scale_p = TRUE, scale_e = FALSE, logxform_e = FALSE, exposure_levels = c(1, 2) )
```
## 3. Print tidied summary stats for the model ( #2, full model)

```{r}
ucr_ucl$models[[2]]$tidied
```
What to expect: 

Since URXUCL is now treated as a factor with two levels (baseline = 1, comparison = 2), the returned coefficient
indicates the mean difference in urinary creatinine (in SD units) between Chlamydia‐positive vs. Chlamydia‐negative, after adjusting for age, sex, race/ethnicity, etc.

# 8.  Disconnect from the Database

```{r}
disconnect_pewas_data(con)
```


# 9.  Summary and Next Steps: You’ve seen how to perform four different phenotype–exposure associations:

    1.  Continuous P ~ Continuous E (no log for E).
    
    2.  Continuous P ~ Continuous E (no log for E).
    
    3.  Continuous P ~ Continuous E with custom covariate set (cycle + BMI, plus default).
    
    4.  Continuous P ~ Categorical E (two‐level exposure). 
      • You’ve also learned how to: 
      • Inspect built-in covariate scenarios
        (adjustment_models). 
        • Create your own custom scenario by
        appending rows. 
        • Extract tidy regression output and R² differences.

    • Check out the web‐based PE Atlas for a broader overview of all associations
    http://apps.chiragjpgroup.org/pe_atlas/
    
