---
title: "Select part 2"
author: "Chirag Patel"
date: "2023-03-12"
output: html_document
---

- figure out what is categorical 
- which to add, which to remove


```{r libraries}
library(tidyverse)
library(DBI)
library(survey)
library(dplyr)
```


```{r db}
dbname <- "nhanes_122322.sqlite"
con  <- dbConnect(RSQLite::SQLite(), dbname)
```

```{r select part 2}
variable_information <- tbl(con, "variable_names_epcf")
table_information <- tbl(con, "table_names_epcf") # c - control; e - exposure, f-filter; p - phenotype
variable_information_joined <- variable_information |> left_join(table_information |> select(Data.File.Name , epcf), by="Data.File.Name")
selected_variables <- read_csv("./select_variables_1.csv") |> filter(select == 1)
variable_information_selected <- variable_information_joined |> collect() |> filter(Variable.Name %in% selected_variables$Variable.Name)
```

#  refer to bp_smq_paq_update.Rmd
- add variables created there in the selected list

```{r}
updated_bpx_variables <- variable_information_joined |> filter(Variable.Name == "MSYSTOLIC" | Variable.Name == "MDIASTOLIC")
updated_smq_variables <- variable_information_joined |> filter(Variable.Name == "SMQ_current_ever_never"  | Variable.Name ==  "SMQ_how_long_since_quit_years") 
pa_names_ad <- c("walk_min_per_day", 
               "PAQ_walk_days",
               "PAQ_walk_days_per_week", 
               "PAQ_walk_met",
               "PAQ_home_week", 
               "PAQ_home_met",
               "PAQ_work_met",
               "PAQ_strengthening_days",
               "PAQ_strengthening_week", 
               "PAQ_strengthening_met" ,
               "PAQ_sedentary_activity_min_per_day",
               "PAQ_leisure_met" , 
               "PAQ_total_met")

pa_names_ej <- c("PAQ_work_vigorous_week",
                 "PAQ_work_moderate_week",
                 "PAQ_walk_week",
                 "PAQ_recreation_vigorous_week",
                 "PAQ_recreation_moderate_week",
                 "PAQ_work_vigorous_met" ,
                 "PAQ_work_moderate_met",
                 "PAQ_walk_met",
                 "PAQ_recreation_vigorous_met" ,
                 "PAQ_recreation_moderate_met",
                 "PA_sedentary_activity_min_per_day",
                 "PAQ_total_met")
updated_paq_variables <- variable_information_joined |> filter(Variable.Name %in% c(pa_names_ad, pa_names_ej))

updated_variables <- rbind(updated_bpx_variables |> collect(),updated_smq_variables |> collect(), updated_paq_variables |> collect()) 

# remove the SMQ 
# remove the PA variables
smq_tables <- variable_information_joined |> filter(Variable.Name == "SMQ_current_ever_never") |> pull(Data.File.Name) |> unique()
pa_tables <- variable_information_joined |> filter(Variable.Name == 'PAQ_total_met' ) |> pull(Data.File.Name) |> unique()
variable_information_selected <- variable_information_selected |> filter(!(Data.File.Name %in% smq_tables))
variable_information_selected <- variable_information_selected |> filter(!(Data.File.Name %in% pa_tables))
variable_information_selected <- variable_information_selected |> rbind(updated_variables)

```


filter out variables that are not in the database
```{r}
in_db <- dbListTables(con)
variable_information_selected <- variable_information_selected |> filter(Data.File.Name %in% in_db)
```

```{r}
remove_vars <- c("DXASPNST", "BMDSADCM", "BPXDAR", "BMXSAD1", "BMXSAD2", "BMXSAD3", "BMXSAD4", "HIQ065" , "BPXDI1", "BPXDI2", "BPXDI3", "BPXDI4", "BPXSY1", "BPXSY2", "BPXSY3", "BPXSAR", "LBDSELSI", "LBDSY3", "LBDSY4")
variable_information_selected <- variable_information_selected |> filter(!(Variable.Name %in% remove_vars))
variable_information_selected <- variable_information_selected |> filter(!(Variable.Name == 'URXUCR' & epcf == 'e'))
```


## bookkeeping categorical: see categorical_levels_table.R

```{r}
e_variable_levels <- tbl(con, "e_variable_levels")
```


# Housing characteristics
# HOQ065: --> 1: owned, 2: rented, 3: other arrangement, 7: refused


#Income
# IND310 - categorical variable that has 77 and 99 as refused, dont know, missing
# INQ300 - categorical variable of family savings > 20k  -> 1 = yes, 2=no
# Income from DEMO

# HPV
# HPV: 
# keep LBDHPCR,ORXHPV,LBXHP2C  remove everything else
# 1: Positive, 2: Negative

```{r}
hpv_variable_names <- c(
  'LBDHPCR',
  'LBDRPCR',
  'ORXHPV',
  'LBXHP2C',
  'LBX06', # - HPV 06 result
  'LBX11', # - HPV 11 result
  'LBX16', # - HPV 16 result
  'LBX18' # - HPV 18 result
)
hpv_rows <- variable_information_selected |> filter(Variable.Name %in% hpv_variable_names)
# remove the tables 
hpv_tables <- variable_information_selected |> filter(Variable.Name %in% hpv_variable_names) %>% pull(Data.File.Name)
variable_information_selected <- variable_information_selected |> filter(!(Data.File.Name %in% hpv_tables))
variable_information_selected <- variable_information_selected |> rbind(hpv_rows)
  
```

```{r}
to_remove <- c('LBDSY4', 'LBXBVPH', 'LBXHIV1', 'LBXHIV2', 'LBXHNAT')
variable_information_selected <- variable_information_selected |> filter(!(Variable.Name %in% to_remove))
```

# S. Aureus
# LBXMS1
```{r}
saur_table_names <- variable_information_selected |> filter(Variable.Name == 'LBXMS1') |> pull(Data.File.Name) |> unique()
saur_rows <- variable_information_selected |> filter(Variable.Name == 'LBXMS1')
variable_information_selected <- variable_information_selected |> filter(!(Data.File.Name %in% saur_table_names))
variable_information_selected <- variable_information_selected |> rbind(saur_rows)

```

remove SS for now
```{r}
variable_information_selected <- variable_information_selected |> filter(!grepl('^SS', Data.File.Name))
```

```{r}
variable_information_selected |> filter(epcf == 'p') |> arrange(Variable.Name) |> write_csv(file="select_pheno_variables_2.csv")
variable_information_selected |> filter(epcf == 'e') |> arrange(Variable.Name) |> write_csv(file="select_expo_variables_2.csv")
```

## may want to analyze BNP

# Education
# DMDEDUC2 - new version
# DMDEDUC3 - for kids

