---
title: "Oral Health uBiome"
author: "Chirag Patel"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Oral Health uBiome-wide Associations

- "TOOTH_DECAY_OHAROCDT", "DENTURE_OHAROCDE","ORAL_HYGIENE_OHAROCOH","GUM_DISEASE_OHAROCGP"

## Model Specification Example

- Tooth Decay ~ genus_i + age + sex + poverty + education + race

```{r}
library(tidyverse)
devtools::load_all("..")
path_to_db <- '../db/nhanes_112824.sqlite' ## path to the sqlite 
con <- DBI::dbConnect(RSQLite::SQLite(), dbname=path_to_db) 
variable_names <- tbl(con, "variable_names_epcf") 
```

```{r}

oral_d_tidied <- rbind(
  read_rds("./TOOTH_DECAY_OHAROCDT.rds")$pe_tidied,
  read_rds("./TOOTH_DECAY_OHAROCDT_RB.rds")$pe_tidied,
  read_rds("./GUM_DISEASE_OHAROCGP.rds")$pe_tidied,
  read_rds("./GUM_DISEASE_OHAROCGP_RB.rds")$pe_tidied,
  read_rds("./ORAL_HYGIENE_OHAROCOH.rds")$pe_tidied,
  read_rds("./ORAL_HYGIENE_OHAROCOH_RB.rds")$pe_tidied,
  read_rds("./DENTURE_OHAROCDE.rds")$pe_tidied,
  read_rds("./DENTURE_OHAROCDE_RB.rds")$pe_tidied
)

#oral_d_tidied <- rbind(tooth_decay$pe_tidied,
#                       gum_disease$pe_tidied,
#                       oral_hygiene$pe_tidied,
#                       denture$pe_tidied
#                       )

```

```{r}
oral_d_tidied_2 <- oral_d_tidied |> filter(model_number == 2, term == "expo")
oral_d_tidied_2 <- oral_d_tidied_2 |> left_join(variable_names |> filter(Begin.Year == 2009) |> select(Variable.Name, Variable.Description) |> collect(),by=c("exposure"="Variable.Name")) |> rename(exposure_description = Variable.Description)
```


What is the SD and the total non-zero count of the uBiome variables?
```{r}
rsv_relative <- tbl(con, 'DADA2RSV_GENUS_RELATIVE_F') |> collect() |> rbind(tbl(con, 'DADA2RSV_GENUS_RELATIVE_G') |> collect())
rb_relative <- tbl(con, 'DADA2RB_GENUS_RELATIVE_F') |> collect() |> rbind(tbl(con, 'DADA2RB_GENUS_RELATIVE_G') |> collect())

rsv_sd <- rsv_relative |> summarise(across(where(is.numeric), list(sd = sd), .names = "{.col}")) |> 
  pivot_longer(cols = everything())

rsv_non_zero <- rsv_relative |> summarise(across(where(is.numeric), list(non_z = ~sum(.x>0)), .names = "{.col}")) |> 
  pivot_longer(cols = everything())

rb_non_zero <- rb_relative |> summarise(across(where(is.numeric), list(non_z = ~sum(.x>0)), .names = "{.col}")) |> 
  pivot_longer(cols = everything())

sd_non_zero <- rsv_sd |> rename(SD=value) |> left_join(rsv_non_zero, by="name") |> rename(non_zero_cnt = value)

```

```{r}
oral_d_tidied_2 <- oral_d_tidied_2 |> left_join(rsv_non_zero |> rbind(rb_non_zero), by=c("exposure"="name")) 
oral_d_tidied_2 <- oral_d_tidied_2 |> rename(non_zero_cnt=value)
```

# Clinical Variables and uBiome Genus with at least X non-zero values
- set X = 1000, approx 10%
```{r}
oral_d_tidied_2 <- oral_d_tidied_2 |> filter(non_zero_cnt >= 1000)
oral_d_tidied_2 |> group_by(phenotype) |> count()
oral_d_tidied_2 |> group_by(exposure_description) |> count()
oral_d_tidied_2 <- oral_d_tidied_2 |> mutate(p.value.fdr = p.adjust(p.value, "BH"))
```


# Volcano plot of the oral associations

```{r}
p <- ggplot(oral_d_tidied_2, aes(estimate, I(-log10(p.value.fdr)), color=phenotype ))
p <- p + geom_point() +scale_x_continuous(limits=c(-.75, .75)) + theme_bw() + geom_hline(yintercept=-log10(.05))
p <- p + theme(legend.position="bottom")
p
```

# Tooth Decay

```{r}
knitr::kable(oral_d_tidied_2 |> filter(phenotype == 'TOOTH_DECAY_OHAROCDT', p.value.fdr < 0.05) |> select(exposure_description, estimate, p.value.fdr) |> arrange(p.value.fdr))
```

# Gum Disease

```{r}
knitr::kable(oral_d_tidied_2 |> filter(phenotype == 'GUM_DISEASE_OHAROCGP', p.value.fdr < 0.05) |> select(exposure_description, estimate, p.value.fdr) |> arrange(p.value.fdr))
```

# Oral Hygiene

```{r}
knitr::kable(oral_d_tidied_2 |> filter(phenotype == 'ORAL_HYGIENE_OHAROCOH', p.value.fdr < 0.05) |> select(exposure_description, estimate, p.value.fdr) |> arrange(p.value.fdr))
```

# Dentures 

```{r}
knitr::kable(oral_d_tidied_2 |> filter(phenotype == 'DENTURE_OHAROCDE', p.value.fdr < 0.05) |> select(exposure_description, estimate, p.value.fdr) |> arrange(p.value.fdr))
```
